Network Configuration Synthesis with Abstract Topologies

### 2016-11-12 paper/abstraction/paper.tex
stashing changes

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. On its own, this abstraction loses a lot of information about the underlying structure of the concrete network, making it difficult to reason precisely about fault-tolerance.
%which would hinder a useful analysis of its fault-tolerance.
For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce several additional concepts. The first is the idea of topology hierarchy captured by P and Q, which say that nodes in the tor and aggregator roles are grouped into a number of smaller pods.
The second is a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of edges (and nodes) that may appear in any given concrete network. Operators can capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ say that, within any pod P, the number of outgoing edges from a node in the $TG$ role to a node in the $AG$ role is equal to the number of nodes in the $AG$. Similarly, the number of outgoing edges from $AG$ to $TG$ is equal to the number of nodes in the $TG$ role. These constraints capture the fact that, within any given pod, the global aggregators and ToRs are in a full mesh. Furthermore, the constraints $AG = AL$ and $AG \leq S$ ensure that, within pods P and Q, the $AG$ and $TG$ roles have the same number of routers, which is less than the number of routers in the spine role $S$.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Similar constraints appear for the local aggregator role. The constraint $2 \leq S \leq 4$ makes explicit the possibility for growth, for example by growing the network from that of Figure~\ref{fig:example} to that of Figure~\ref{fig:example2}. In general, we need not bound the number of spine routers to admit more concrete topologies at the expense of analysis precision. Finally, we include the constraints $(S \mod AG) = 0$, and $(S \mod AL) = 0$ simply to show that constraints do not have to inequalities. In general operators can write constraints using logical formula for any theory supported by modern SMT solvers.

\ryan{mention mincut}

@@

The rule \infrule{I-mincut} can be thought of as an assertion about the fault tolerance between two roles. The rule applies to an edge that has been annotated with a $mincut(X)$ statement, which means that there are at least $X$ disjoint paths between any node in role $m$ and any node in role $n$. Annotations of this form are useful for a "one big switch" \ryan{cite} type of abstraction where part of a complex network is abstracted into a single node. For example, to model an ISP backbone network the operator might abstract the network into separate geograph regions with two abstract roles per region. One role for the border routers of the network, and another role for the core of the network. Mincut annotations could be used to describe the fault tolerance both within region cores as well as across regions.

>>>>

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. On its own, this abstraction loses a lot of information about the underlying structure of the concrete network, making it difficult to reason precisely about fault-tolerance.
%which would hinder a useful analysis of its fault-tolerance.
For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce several additional concepts. The first is the idea of a topology hierarchy captured by P and Q, which say that nodes in the tor and aggregator roles are grouped into a number of smaller pods.
The second is a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of edges (and nodes) that may appear in any given concrete network. Operators can capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ say that, within any pod P, the number of outgoing edges from a node in the $TG$ role to a node in the $AG$ role is equal to the number of nodes in the $AG$ role. Similarly, the number of outgoing edges from a node in the $AG$ role to a node in the $TG$ role is equal to the number of nodes in the $TG$ role. These constraints capture the fact that, within any given pod, the global aggregators and ToRs are in a full mesh. Furthermore, the constraints $AG = AL$ and $AG \leq S$ ensure that, within pods P and Q, the $AG$ and $TG$ roles have the same number of routers, which is less than the number of routers in the spine role $S$.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Similar constraints appear for the local aggregator role. The constraint $2 \leq S \leq 4$ makes explicit the possibility for growth, for example by growing the network from that of Figure~\ref{fig:example} to that of Figure~\ref{fig:example2}. In general, we need not bound the number of spine routers to admit more concrete topologies, potentially at the expense of analysis precision. Finally, we include the constraints $(S \mod AG) = 0$, and $(S \mod AL) = 0$ simply to show that constraints do not have to be inequalities. In general operators can write constraints using logical formula from any theory supported by modern SMT solvers.

A final type of abstraction is the $mincut(1)$ constraint between the spine role S and N[1--2]. The mincut constraint says that any node in the spine role has at least one path to any node in the neighbor N1 (and N2) role. Annotations of this form are often useful for a "one big switch" \ryan{cite} type of abstraction where part of a complex network is abstracted into a single node. For example, to model an ISP backbone network the operator might abstract the network into separate geographic regions with two abstract roles per region -- one role for the border routers of the network, and another role for the core of the network. Mincut annotations could be used to describe the fault tolerance both within region cores as well as across regions.

@@

The rule \infrule{I-mincut} can be thought of as an assertion about the fault tolerance between two roles. The rule applies to an edge that has been annotated with a $mincut(X)$ statement, which means that there are at least $X$ disjoint paths between any node in role $m$ and any node in role $n$.
